\documentclass[11pt]{article}

\usepackage[margin=0.9in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{ragged2e}
\usepackage{setspace}
\usepackage[table]{xcolor}
\usepackage[natbibapa]{apacite}

\definecolor{PrimaryRed}{HTML}{8B1E1E}
\definecolor{SoftBlue}{HTML}{DDEEFF}
\definecolor{SoftGray}{HTML}{F4F4F4}

\hypersetup{colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black}
\setlength{\parindent}{0pt}
\setstretch{1.08}
\bibliographystyle{apacite}
\newcommand{\reviewsection}[1]{\vspace{0.07in}\noindent{\normalsize\bfseries\color{PrimaryRed}#1}\par\vspace{0.02in}}

\begin{document}

\hypersetup{pageanchor=false}
\begin{titlepage}
\newgeometry{margin=0.7in}
\thispagestyle{empty}

\begin{center}
{\large University of Rochester\\}
{\normalsize Warner School of Education\\[0.15in]}

{\color{PrimaryRed}\rule{\textwidth}{1.4pt}}\\[0.12in]
{\Huge\bfseries Who Gets to Name,\\[0.07in]
Interpret, and Design?}\\[0.15in]
{\huge Understanding Developmental Difference\\[0.07in]
Through Labels, Lived Experience,\\[0.07in]
and Educational Environments}\\[0.18in]
{\Large EDE 448 --- Module 1 Journal}\\[0.12in]

\vfill

{\color{PrimaryRed}\rule{0.78\textwidth}{0.7pt}}\\[0.22in]

\colorbox{SoftBlue}{%
\parbox{0.88\textwidth}{%
\centering\small\vspace{0.02in}
\textbf{From Deficit Interpretation to Access Design}\\[0.05in]
A reflective analysis of how diagnosis, scientific discourse, neurodiversity, media, schooling, family knowledge, lived experience, and classroom design shape what the world believes about developmental difference.
}%
}
\end{center}

\vfill

\begin{center}
\renewcommand{\arraystretch}{1.18}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{>{\bfseries\RaggedLeft}p{1.65in} p{5.0in}}
Student & Piter Zacari Garcia Bautista \\
Assignment & Module 1 Journal \\
Course & EDE 448 --- Communication and Positive Behavioral Supports for Students with Autism and Other Complex Needs \\
Essential Question & How does the world come to understand developmental differences? \\
Source Conversation & Full Module 1 reading and media set, with selected sources examined in depth and the wider set used to validate, extend, or challenge the analysis \\
Interpretive Lenses & Puzzle Plan and EQUITAS \\
Submission Date & July 2026 \\
\end{tabular}
\end{center}

\vfill

\begin{center}
\begin{minipage}{0.92\textwidth}
\centering\itshape
A label can open access, but it becomes harmful when adults allow it to replace curiosity.\\
{\color{PrimaryRed}\rule{4in}{0.8pt}}
\end{minipage}
\end{center}

\vfill

\begin{center}
\begin{minipage}{0.95\textwidth}
\small
This journal argues that developmental difference is simultaneously neurological, lived, relational, cultural, historical, and socially interpreted. Rather than separating course ideas from experience, each section uses my lived experience and K--5 technology teaching placement to interpret the full Module 1 source conversation. The sources do not always agree, and that disagreement matters: diagnostic and educator resources can open access, while critical, historical, neurodiversity, and lived-experience sources warn against allowing categories, interventions, or success narratives to erase the person.
\end{minipage}
\end{center}

\restoregeometry
\end{titlepage}
\hypersetup{pageanchor=true}
\pagenumbering{arabic}

\reviewsection{What I Noticed Before I Had the Language}

Before these readings, I already knew what it felt like to be interpreted through fragments. As a neurodivergent person, I have lived through institutions that noticed missed deadlines, uneven performance, communication differences, or difficulty navigating an inaccessible process before they noticed effort, context, intelligence, or persistence. That history shapes how I now look at students. When a child hesitates, repeats a question, moves away from a group, keeps a familiar device close, or needs more time, I do not see a complete explanation. I see a clue.

That is the most important shift Module 1 helped me name. The world does not encounter developmental difference directly. It encounters behavior, language, stories, reports, family knowledge, diagnostic criteria, media images, school routines, scientific claims, and policy categories. Public-health and general educational resources give the world recognizable definitions of autism \citep{cdcASD,simon2016asd,readingRocketsASD}. Those definitions can be useful, but \citet{orsiniDavidson2013critical} show what is lost when statistical urgency turns varied autistic lives into one easily understood crisis. Their warning validated something I have felt personally: a system can collect accurate fragments and still produce an inaccurate person.

The historical reading deepened that point. \citet{harvey2018story} presents autism as a changing story rather than a timeless object that institutions simply discovered. The categories, explanations, and expectations surrounding autism have changed as medicine, schools, families, researchers, media, and autistic communities have changed. My central learning is therefore personal as well as academic: developmental difference is neurological and real, but the meanings attached to it are historical, relational, cultural, and political. No single system gets to explain the whole person.

\reviewsection{When a Label Opens a Door---and When It Starts Closing One}

In my elementary technology teaching placement, I repeatedly saw students whose competence was easy to miss because of what surrounded the task. A learner might understand the coding concept but struggle with a long verbal direction, an unfamiliar login path, reading load, noise, peer control of the device, or uncertainty about what finished work should look like. When I modeled the exact tool path, used a visible first/next/last sequence, broke the task into smaller steps, or gave additional processing time, participation changed. The academic idea had not suddenly entered the student's mind. The environment had stopped hiding what was already there.

That experience is what made \citet{cole2023autism} meaningful to me. Their clinical and educational language can give families and professionals a shared vocabulary for social communication, sensory processing, routines, focused interests, and support needs. The CDC and broader school-centered overview similarly validate the practical value of diagnostic literacy: identification can open eligibility, services, and coordinated support \citep{cdcASD,simon2016asd}. At the same time, all three sources acknowledge heterogeneity. A diagnostic label cannot tell me what a particular learner understands, fears, enjoys, prefers, or needs in a specific moment. It cannot tell me whether the barrier is sensory, linguistic, technological, social, emotional, or simply a confusing set of directions.

The wider source set also keeps this claim from becoming too comfortable. \citet{nadesan2013genetics} asks how genetic knowledge becomes entangled with profit, risk, policy, and judgments about which lives are desirable. \citet{orsiniDavidson2013critical} likewise challenge epidemic language and the false choice between treating autism only as a disorder or only as a difference. These sources do not ask educators to discard medical knowledge. They ask us to notice what medical and scientific narratives do once they enter schools, families, funding systems, and public imagination.

This is why labels can become dangerous when adults let them replace curiosity. They influence what people expect, which behaviors are interpreted as meaningful, and which students are assumed capable before they have a chance to show it. Diagnosis is one source of evidence, but it is not the whole story. The challenge is not to reject diagnostic knowledge; it is to prevent that knowledge from becoming the only knowledge adults treat as credible.

\reviewsection{Neurodiversity Became Real to Me Through Both Identity and Teaching}

\citet{ortega2013cerebralizing} connected with something I have felt for years: labels are not only clinical. They can become part of identity, community, resistance, and self-understanding. As someone who received an ADHD diagnosis later in life, I know the relief of finding language that explains patterns previously interpreted as laziness, inconsistency, or failure. A name can reduce shame. It can help a person reinterpret the past and ask for support without believing they are morally defective.

But Ortega also warns that an affirming brain-based identity can become reductive if it suggests that a person is only a brain or that everyone with the same label shares one political, cultural, or personal experience. That tension appears in classrooms too. A teacher can move away from deficit language and still create a new stereotype: the autistic child who is always visual, the student with ADHD who is always energetic, or the neurodivergent learner who is automatically gifted in one narrow way.

My students keep reminding me that difference does not arrive in neat categories. In one activity, a student needed a familiar laptop nearby while also building with blocks. The device could have been interpreted as distraction or refusal. Instead, I treated it as part of the student's regulation and allowed both forms of participation to coexist. The student remained connected to the activity.

\citet{conn2016play} helped me see why that moment mattered beyond accommodation. Conn challenges the familiar myth that autistic children do not truly play and questions adult-directed approaches that measure whether a child performs the outward signs of normative play. The reading directly complicates any intervention that values visible compliance more than internal engagement, enjoyment, or relationship. It also reports that autistic children may sustain more interaction when partners are less directive and allow more time. That does not mean adults withdraw support; it means support must leave room for the learner's authorship.

The Autism Speaks reading list is useful as a pathway to family-facing and public-facing resources, but even a resource list carries assumptions about whose books, voices, and purposes count \citep{autismSpeaksReads}. It therefore needs to be paired with autistic-authored, critical, and neurodiversity-affirming work. For me, neurodiversity is most useful when it expands whose knowledge counts and protects variation within the spectrum. It should make us more curious, not more certain.

\reviewsection{What the Media Helped Me Recognize About Representation}

Watching \textit{Autism: Insight from Inside} reminded me of how strongly public understanding depends on stories \citep{infobase2016inside}. I have experienced the power of being seen through the wrong story: the unreliable student, the overly emotional person, or the one who should simply try harder. I also know the relief of encountering a story that makes room for complexity rather than forcing a person into a deficit narrative.

The film validates several claims in the readings. Its participants describe labels as both access points and sources of stigma, and they insist that people with the same diagnosis can have very different sensory experiences, communication profiles, interests, and support needs. At the same time, parts of the film complicate my own argument. Its strong emphasis on work skills, independence, productivity, and ``good outcomes'' can unintentionally imply that acceptance must be earned through economic usefulness. Building strengths is important, but a person's dignity cannot depend on becoming employable, independent, or exceptional.

The broader media menu makes this tension even clearer. \textit{Autism: The Musical} shows children creating, performing, relating, and surprising adults, which strongly validates the claim that environments can reveal competence \citep{regan2007musical}. Yet its older dialogue also includes pressure to ``blend in'' and family movement through denial, fixing, and acceptance. The course warning that the film is nearly two decades old matters because representation can affirm possibility while still carrying normalization assumptions from its time.

\textit{Best Kept Secret} extends the question beyond childhood by following autistic students in Newark as they approach the end of public-school services \citep{buck2013best}. It supports my claim that disability is socially interpreted, but it also forces a harder question: what does school-based inclusion mean if students encounter a service cliff afterward? Its attention to race, economic inequality, and adult transition strengthens the intersectional argument that access is never distributed evenly.

The course also offers \textit{Talk to Me} and \textit{Unspoken: The Dynamic Life of an Autistic Teenager} \citep{talkToMe2007,unspoken2018}. Because the licensed versions were not locally transcribed, I do not assign them detailed claims that I cannot verify. Still, their placement in this module matters: communication and adolescence are not side topics but central sites where adults decide whose expression is recognizable and whose life is considered knowable. The American Autism Association podcast pathway further broadens the media field through parent, educator, sensory, faith-based, Black autism, and first-person perspectives \citep{aaa2022podcasts}. That variety is useful, but a list of voices is not automatically balanced; educators still need to ask whether autistic people are speaking for themselves or are mainly being spoken about.

Media can therefore support, extend, and oppose the same claim at once. It can widen public understanding by making hidden experience visible. It can also create tragedy narratives, inspiration narratives, productivity narratives, or one ``successful'' autistic person who becomes a measuring stick for everyone else.

\reviewsection{What Students Taught Me About Behavior, Communication, and Design}

The National Education Association guide became most useful when I connected it to moments from my classroom \citep{nea2020guide}. During technology, coding, and robotics activities, I saw how movement, repetition, avoidance, silence, leaving a group, or refusing a material could be interpreted as defiance when the actual issue was overload, uncertainty, lack of control, or inaccessible directions. I also saw how quickly behavior changed when the environment changed.

When I placed only the needed materials on the table, modeled the first step, reduced the number of verbal directions, clarified peer roles, or offered a quieter starting point, students often re-entered the activity without a lecture, consequence, or reward. The practical educator sources validate those moves. Reading Rockets connects autism knowledge to literacy and communication access; the IRIS educator module and autism educator guide emphasize structured, flexible support; and \citet{denningMoody2018} argue for proactively changing classroom practices rather than waiting for a student to fail \citep{readingRocketsASD,iris2016asd}. Denning and Moody are especially helpful because they name a problem I saw repeatedly: classrooms are often organized around a presumed typical learner, and then individual students are treated as disruptions when they need the environment to change.

The inclusion sources validate the same point from another direction. \citet{bock2016inclusion} distinguish presence from meaningful participation and emphasize that teacher knowledge, attitudes, training, and administrative support affect whether inclusion becomes real. \citet{smith2011inclusion} likewise treats inclusion as planned, supported work rather than a hopeful placement decision. These sources strengthen my claim that inclusion requires structure and accountability. At the same time, their intervention-oriented language must be read alongside Conn and neurodiversity scholarship so that ``support'' does not quietly become training a student to appear more typical.

The strongest question I can ask is not, \textit{``How do I stop this behavior?''} It is, \textit{``What might this student be communicating, and what barrier is present?''} That change in language changes the intervention. It moves me away from controlling the student and toward redesigning the conditions.

This also changed how I understand independence. Support should not mean doing the task for a student or making every decision on the student's behalf. The strongest scaffolds make the goal visible and the entry point accessible while preserving opportunities for choice, productive struggle, communication, revision, joy, and repair. A visual sequence, partner role, sensory option, communication board, model, or quiet reset should reveal competence rather than replace it.

\reviewsection{What Counts as a Good Outcome?}

My placement taught me to celebrate when students could enter an activity, communicate a need, recover from frustration, contribute an idea, or remain connected in their own way. Module 1 pushed me to ask whether schools sometimes define a ``good outcome'' too narrowly: task completion, quiet behavior, independence from support, employment, or the ability to resemble a typical peer.

The transition reading expands support beyond the immediate classroom. \citet{stuartCollins2019transition} connect visual supports, technology, self-determination, and transition planning to adult and community participation. Their work validates the importance of preparing for the future, but it also changes what preparation should mean. Future planning should begin with the student's goals, communication, preferences, relationships, and desired community life---not only with adult assumptions about productivity.

\textit{Best Kept Secret} makes that institutional problem visible through students nearing the end of public education \citep{buck2013best}. The film and transition reading together oppose a classroom-only understanding of access. A school has not fully supported agency if the student leaves with no meaningful pathway into adult services, relationships, community, continued learning, or self-directed life. Independence can be valuable, but interdependence is not failure. A good outcome is not a life without support; it is a life in which support does not erase authorship.

\reviewsection{Why Credibility, Culture, and Identity Cannot Be Separated}

My lived experience makes it impossible for me to think about developmental difference as an isolated variable. I am Latino, bilingual, neurodivergent, chronically ill, and shaped by trauma. Different institutions have interpreted the same traits differently depending on which part of me they noticed first. Sometimes persistence was seen as intensity. Sometimes difficulty with an inaccessible process was treated as lack of responsibility. Sometimes communication style mattered more than the content of what I was saying.

Students also move through the world with multiple identities. An autistic learner may be multilingual, racialized, gender-diverse, economically marginalized, physically disabled, or part of a family and culture with its own language for disability and support. Those identities affect who is noticed, who is believed, who receives an evaluation, which behavior is punished, and what services are available. \citet{harvey2018school} reinforces that school is not a neutral setting; its routines and expectations reward some bodyminds while pressuring others to mask. \textit{Best Kept Secret} adds the material consequences of race, poverty, geography, and transition access to that analysis \citep{buck2013best}.

This is why lived experience cannot be treated as an optional add-on after the ``real'' academic analysis. Professional knowledge is incomplete without it. Family knowledge becomes incomplete only when institutions decide not to listen. Student behavior is incomplete evidence when adults refuse to consider context. Critical autism studies, historical accounts, genetic discourse, educator guides, family resources, and media narratives all reveal different pieces; they also carry different power to define the person \citep{orsiniDavidson2013critical,nadesan2013genetics,harvey2018story,autismSpeaksReads}.

A culturally responsive understanding therefore requires more than adding identity words to a paragraph. It requires asking whose observations become data, whose language becomes diagnostic, whose story becomes representative, and whose goals shape support.

\reviewsection{Designing Better Evidence Through the Puzzle Plan and EQUITAS}

The full Module 1 source set comes together for me in the Puzzle Plan. Diagnostic systems, school systems, scientific systems, media systems, and technology systems all create evidence about people. A shallow system records the label, error, delay, referral, absence, behavior, test score, or failed task. A richer system also records task demands, sensory conditions, communication options, language, tool access, peer dynamics, prior knowledge, student preference, family knowledge, historical context, cultural identity, and what changed when support was offered.

I have seen the difference in real time. When a student struggled with a technology task, the shallow evidence was that the student did not complete it. The richer evidence included a confusing login sequence, too much spoken information, peer competition for the device, noise, unclear visual expectations, and insufficient processing time. Once those conditions changed, the student's participation changed too.

The source disagreements are part of that richer evidence. Diagnostic and public-health sources help identify patterns and access services. Practical educator texts identify concrete supports. Inclusion and transition scholarship demand planning and accountability. Neurodiversity and critical autism studies challenge who controls the meaning of autism. Historical and genetic analyses expose the power inside scientific stories. Play scholarship resists adult control masquerading as development. Media reveals possibility while also carrying assumptions about tragedy, inspiration, productivity, normality, and whose voice gets centered.

The quality of the evidence shapes the quality of the response. If the evidence is deficit-only, the intervention will usually ask the student to become easier for the system. If the evidence includes context, disagreement, voice, and agency, educators can redesign the system to become more responsive to the student.

Through the EQUITAS lens, I now ask whether classroom structures provide equitable and accountable pathways for students to access learning, communicate needs, experience joy, demonstrate competence, build relationships, and exercise agency without erasing difference in order to belong. That question is not separate from the readings or media. It is what the full Module 1 conversation becomes when it meets a real student, a real classroom, and a real life.

\reviewsection{Conclusion}

The world comes to understand developmental differences through labels, statistics, scientific explanations, stories, relationships, institutions, environments, and the voices it chooses to treat as credible. My own life and teaching have shown me that a person can be described very differently depending on whether a system notices only the difficulty or also notices the conditions surrounding it.

The Module 1 sources support parts of that claim, extend other parts, and challenge me where my thinking could become too simple. Clinical and educator sources validate the need for shared language and concrete support. Critical, historical, genetic, neurodiversity, and play scholarship warn that help can reproduce harm when it treats normality, compliance, productivity, or adult interpretation as the goal. The media shows both the power and danger of representation: stories can reveal humanity, but they can also turn one life into a standard for everyone else.

A label can open access, but it becomes harmful when adults allow it to replace curiosity. In my future classroom, I want to combine diagnostic information with direct observation, student communication, family knowledge, sensory and environmental information, peer dynamics, historical and cultural context, lived experience, and collaboration with specialists. I want support to reveal competence rather than conceal agency.

The goal is not to make every learner communicate, move, focus, play, work, or participate in the same way. The goal is to create a classroom where students with autism and other complex support needs can communicate, make choices, experience joy, demonstrate competence, belong, and be believed. Understanding developmental difference responsibly requires more than naming a condition. It requires staying curious about the person, examining the power inside our sources, collecting richer evidence, and changing the environment before deciding that the learner is the problem.

\newpage
\small
\bibliography{references}

\end{document}
